---
title: "Presenting to a Technical Audience"
author: "Kirin Patel"
date: "2/24/2020"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r import libraries, include=FALSE, echo=FALSE}
library(ggplot2)
require(plyr)
```

## Presenting to a Technical Audience

This assignment focuses on using R to process data that can then be used to
generate a PowerPoint presentation that can be presented to a technical or
topic area familiar audience.

## Getting Started

Before analysis can be conducted, a dataset needs to be obatained. For this
assignment, I will be using two datasets from
[FRED](https://fred.stlouisfed.org). The first is the
[Retail Sales: Used Car Dealers](https://fred.stlouisfed.org/series/MRTSSM44112USN)
and the second one is the
[Average Amount Financed for New Car Loans at Finance Companies](https://fred.stlouisfed.org/series/DTCTLVENANM).
Both datasets were found through
[Google's Dataset Search Tool](https://datasetsearch.research.google.com).

```{r load data, echo=FALSE}
used_car_sales_data_url = "https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1168&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=MRTSSM44112USN&scale=left&cosd=1992-01-01&coed=2019-12-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2009-06-01&line_index=1&transformation=lin&vintage_date=2020-02-24&revision_date=2020-02-24&nd=1992-01-01"
used_car_sales_data = read.csv(url(used_car_sales_data_url))
used_car_sales_data$'Year' <- as.numeric(substr(used_car_sales_data$'DATE',1,4))
used_car_sales_data$'Month' <- as.numeric(substr(used_car_sales_data$'DATE',6,7))

average_finance_amount_data_url = "https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1168&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=DTCTLVENANM&scale=left&cosd=2008-03-01&coed=2019-09-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=diamond&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2009-06-01&line_index=1&transformation=lin&vintage_date=2020-02-24&revision_date=2020-02-24&nd=2008-03-01"
average_finance_amount_data = read.csv(url(average_finance_amount_data_url))
# Remove values that are equal to '.' to prevent missing data from being rendered.
average_finance_amount_data <- average_finance_amount_data[average_finance_amount_data$'DTCTLVENANM' != '.',]
average_finance_amount_data$'Year' <- as.numeric(substr(average_finance_amount_data$'DATE',1,4))
average_finance_amount_data$'Month' <- as.numeric(substr(average_finance_amount_data$'DATE',6,7))
```

## Why This Data Should be Analyzed

Looking at the financial history of used car sales and average loan amounced of
financed cars can help inform individuals about the best times for when to buy
or sell a car. This data can help shed light onto when average prices paid for
cars can be maximized or minimized.

## Millions of Dollars Made on Used Cars at Retail Dealers

```{r}
ggplot(data=used_car_sales_data,aes(x=used_car_sales_data$'Year',y=as.numeric(used_car_sales_data$'MRTSSM44112USN'),color=used_car_sales_data$'Month')) +
  geom_point() +
  geom_smooth(method='lm') +
  xlab('Year') +
  ylab('Millions of Dollars')
```

## Reviewing the Data

While only one set of data has been seen, a clear trend can be seen. Months
in the beginning of the year result in higher numbers of sales and higher
profits. Generally, the months of January, February, and March are the
highest selling months per year and have continually resulted in the highest
sales year over year. This trend has only continued to grow.

## Average Financed Amount for New Car Loans

```{r}
ggplot(data=average_finance_amount_data,aes(x=as.numeric(average_finance_amount_data$'Year'),y=as.numeric(average_finance_amount_data$'DTCTLVENANM'),color=average_finance_amount_data$'Month')) +
  geom_point() +
  geom_smooth(method='lm') +
  xlab('Year') +
  ylab('Thousands of Dollars')
```

## Reviewing the Data

Following the trend set in the previous dataset, the average financed amount
for new car loans increases year after year. However, the later months of the
year generally have larger averafe finance amount for new cars.

## Combining Datasets

When looking at both the total amount of sales in USD and average finance
amount of new car loans over the past decade, there is a rough ideal item
presented for when to buy or sell a car. If you are planning on selling a car,
you should do so during the beginning of the year to maximize profits as many
sales are made during this time. However, you should only buy a new car during
the beginning of the year since average finance amounts for new car loans is
lower. Should you be buying a used car, do so towards the end of the year to
maximize any sales.
