---
title: "Reproducibility"
author: "Kirin Patel"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reproducibility is a fundamental portion of data science. Without
reproducibility, the validity of a result cannot be proven, and thus, all
efforts made when conducting experiments are made irrelevant.

For this assignment, I will be reproducting the experiment from David K. Smith,
[Car Datasets](https://www.rpubs.com/dksmith01/cars), with
[this](https://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data)
dataset.


### Installing Libraries

```{r, eval=FALSE}
install.packages("magrittr")
install.packages("plyr")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("grid")
install.packages("gridExtra")
```


### Importing Libraries

```{r}
library(magrittr)
library(plyr)
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
```


### Importing Data

```{r}
data_url = url('https://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.data')
cars = read.csv(data_url, header=FALSE)
```


### Setting Headers

These headers were obtained from [imports-85.names](https://archive.ics.uci.edu/ml/machine-learning-databases/autos/imports-85.names).

```{r}
headers <- c('symboling','normalized-losses','make','fuel-type','aspiration','num-of-doors','body-style','drive-wheels','engine-location','wheel-base','length','width','height','curb-weight','engine-type','num-of-cylinders','engine-size','fuel-system','bore','stroke','compression-ratio','horsepower','peak-rpm','city-mpg','highway-mpg','price')
names(cars) <- headers
```


### Statistics

```{r}
str(cars)
```

```{r}
summary(cars)
```


### Plots

```{r}
# Miles Per Gallon
qplot(cars$'city-mpg',
      xlab = 'Miles Per Gallon (City)',
      ylab = 'Count',
      binwidth = 2, 
      main = 'Frequency Histogram: Miles per Gallon (City)'
)

qplot(cars$'highway-mpg',
      xlab = 'Miles Per Gallon (Highway)',
      ylab = 'Count',
      binwidth = 2, 
      main = 'Frequency Histogram: Miles per Gallon (Highway)'
)

# Number of Cylinders
table(cars$'num-of-cylinders')

# Displacement
qplot(cars$'engine-size',
      xlab = 'Displacement',
      ylab = 'Count',
      binwidth = 20,
      main = 'Frequency Histogram: Displacement'
)

# Weight
qplot(cars$'curb-weight',
      xlab = 'Weight',
      ylab = 'Count',
      binwidth = 200,
      main = 'Frequency Histogram: Weight'
)

# Price
qplot(cars$price,
      xlab = 'Price',
      ylab = 'Count',
      main = 'Frequency Histogram: Price'
)

summary(cars$price)
```


### Bivariable Plots

```{r}
ggplot(data = cars,
       aes(x = cars$'curb-weight', y = cars$'city-mpg')) +
       geom_point() +
       geom_smooth(method='lm') +
       xlab('Weight') +
       ylab('MPG') +
       ggtitle('Weight vs. MPG (City): Entire Sample'
)

ggplot(data = cars,
       aes(x = cars$'make', y = cars$'city-mpg')) +
       geom_boxplot() +
       xlab('Make') +
       ylab('MPG (City)') +
       ggtitle('MPG Comparison by Make'
)
```


### Conclusion

Unfortunately, due to the differences between datasets, I am unable to replicate
some of the more interesting visualizations that David has created. However,
even with this limitation, the same conclusions that David did. As car weight and engine size (displacement) increase, the efficiency of the vehicle
decreases.

For more detailed analysis of car manufaturer trends visit David's
[article](https://www.rpubs.com/dksmith01/cars).
