---
title: "Presenting to a General Audience"
author: "Kirin Patel"
date: "2/17/2020"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r import libraries, include=FALSE, echo=FALSE}
library(ggplot2)
require(plyr)
```

## Presenting to a General Audience

This assignment focuses on using R to process data that can then be used to
generate a PowerPoint presentation that can be presented to a non-tech or
unfamiliar audience.

## Getting Started

Before analysis can be conducted, a dataset needs to be obatained. For this
assignment, I will be using the
[BuildBPS: Facilities and Educational Data for Boston Public Schools](https://data.boston.gov/dataset/buildbps-facilities-and-educational-data-for-boston-public-schools)
which was found through
[Google's Dataset Search Tool](https://datasetsearch.research.google.com).


```{r load data, echo=FALSE}
data_url = "https://data.boston.gov/dataset/b2c5a9d3-609d-49ec-906c-b0e850a8d62a/resource/33c5f44a-3c67-4390-a1d5-1bf018e4728c/download/buildbps.csv"
data = read.csv(url(data_url))
```

## Why This Data Should be Analyzed

Current and previous data collected on educational infastructure provides
insight as to how money is currently being spent, where it is beign
effectively spent, and where spending needs to improve. By investigating this
data, insightful conclusions involving future construction of educational
facilities can be made.

## Yearly Electric Bill of Educational Facilities Based on Year Built

```{r view data}
ggplot(data=data,aes(x=as.numeric(data$'BPS_Year_Built'),y=as.numeric(data$'BPS_Electric_Bill'))) +
  geom_point() +
  xlab('Year Built') +
  ylab('Yearly Electric Bill (Tens of Thousands of Dollars)')
```

## Analysing This Data

Upon first analysis of this data, it appears that educational facilities do not
become more efficient with electricity consumption. However, the data plotted
does not account for school size or type. Both pre-school, elementary, middle,
and highschool facilities were plotted.

## Yearly Electic Bill of Educational Facilities Based on Year Built With Facility Type Highlighted

```{r view data with school type}
ggplot(data=data,aes(x=as.numeric(data$'BPS_Year_Built'),y=as.numeric(data$'BPS_Electric_Bill'),color=data$'SMMA_Typology')) +
  geom_point() +
  xlab('Year Built') +
  ylab('Yearly Electric Bill (Tens of Thousands of Dollars)')
```

## Analysing the Data Again

Now that educational facility types can be connected to facility costs,
analysis on which facilities consume the most electricity can be determined.
Generally, lower education level facilities have lower yearly eltrical bills
than higher level facilities, with the exception of early learning facilities.

## Why Might Higher Level Education Facilities Have Higher Electrical Bills

The most simpilest anwser is that, they are larger facilities. Generally, one
would think that as education level of a facility increases, so does its size.
However, most educational facilities within Boston remain under 500,000 square
feet. Size is not a determining factor for electrical bills.

## Size of Educational Facilities Based on Year Built With Facility Type Highlighted

```{r}
ggplot(data=data,aes(x=as.numeric(data$'BPS_Year_Built'),y=as.numeric(data$'SMMA_Site_SF'),color=data$'SMMA_Typology')) +
  geom_point() +
  xlab('Year Built') +
  ylab('Size of Facility (Square Feet)')
```

## Does Student Population Size Influence Eletrical Bills?

```{r}
ggplot(data=data,aes(x=as.numeric(data$'DOE_Total'),y=as.numeric(data$'BPS_Electric_Bill'),color=data$'SMMA_Typology')) +
  geom_point() +
  xlab('Student Population Size') +
  ylab('Yearly Electric Bill (Tens of Thousands of Dollars)')
```

## No, Student Population Size Does not Influence Eletrical Bills

Based on the data alone, a single field cannot be attributed to the differences
between the eletrical bills of educational facilities. There are likely a
multitude of factors that have differing levels of influence over the
eletrical bills of educational facilities. Further investigation into the
impacts of each factor is required in order to present insightful conclusions
involving future construction of educational facilities can be made.
