---
title: "How to Write a Blog Post"
author: "Kirin Patel"
date: "3/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import libraries, include=FALSE, echo=FALSE}
require(plyr)
library(ggplot2)
library(gridExtra)
```

## Investigating the Affects of Climate Change on Energy Usage in the United States

Within this blog post, I will be investigating the affects of ozone
concentrations within our atmosphere on the energy usage of the United States.
This analysis will be conducted on data collected during the last six months of
2019. These data sets were obtained from the
[EPA](https://www.epa.gov/castnet/castnet-ozone-monitoring)
and the
[EIA](https://www.eia.gov/beta/electricity/gridmonitor/dashboard/electric_overview/US48/US48?src=email).

Before we load any data, understanding what ozone is and what it does is
important. Ozone is within the Earth's atmosphere and reduces the amount of
harmful UV radiation that humans are exposed to, which can lead to sun burns
and skin cancer
([CDC, 2019](https://www.cdc.gov/features/uv-radiation-safety/index.html)).
While this is only one of the risks that UV radiation presents to humans, it
is one that is commonly known and it presents individuals with a reason
potentially increase energy usage. As the sun raises to its highest point at
noon, humans are exposed to more UV radtion, as a result of that, water intake
may be increased and building temperatures may be decreased with air
conditioners, both of which would lead to increased energy usage. While the
specific applications that lead to increased energy usage will not be
investigated, this analysis will see if there is any correlation between
ozone concentrations and energy usage within the United States.

## Loading and Validating the Datasets

Before analysis can begin, the datasets must first be loaded into RStudio.

```{r loading_the_datasets}
power_balance_data_url = 'https://www.eia.gov/realtime_grid/sixMonthFiles/EIA930_BALANCE_2019_Jul_Dec.csv'
power_balance_data = read.csv(url(power_balance_data_url))


# Obtained via the EPA site, specifying 'Aggregate Concentration Data' for the
# last 6 months of 2019 across all sites that they record data from. This data
# is provided as a file rather than a URL due to the inability to directly
# download this dataset.
ozone_daily_max_data = read.csv('./ozone-8-hour-daily-max.csv')
# Obtain date in identical format to power_balance_data.
ozone_daily_max_data$Date <- substr(ozone_daily_max_data$DDATE,1,10)
```

```{r validating_the_power_dataset}
summary(power_balance_data)
```

```{r validating_the_ozone_dataset}
summary(ozone_daily_max_data)
```

The datasets appear to be loading as intended. Now we can analyze them.

## Looking at Power Demand in the United States during the Last six Months of 2019

There are over 60 power authorities for the United States, while each of these
should be investigated and compared with one another, it would take far too
long for the purpose of this exercise. Instead, five of the most major
authorities will be used: CISO, PJM, ERCO, FPL, and MISO.

```{r plot_power_demand_df}
# Generic function to simplify plotting of power demand.
plot_power_demand_df <- function(df) {
  ggplot(data=df,aes(x=df$Data.Date,y=as.numeric(df$Demand..MW.),color=df$Hour.Number)) +
    geom_point() +
    geom_smooth(method='lm') + 
    xlab('Date') +
    ylab('Power Demand (Megawatts)') +
    # Scale date by month to reduce number of labels.
    scale_x_discrete(breaks=c('07/01/2019','08/01/2019','09/01/2019','10/01/2019','11/01/2019','12/01/2019'))
}
```

#### California Independent System Operator Authority
```{r viewing_the_power_dataset_ciso_authority}
df = power_balance_data[power_balance_data$Balancing.Authority == 'CISO',]
ciso_power = plot_power_demand_df(df)
ciso_power
```

#### PJM Authority
```{r viewing_the_power_dataset_pjm_authority}
df = power_balance_data[power_balance_data$Balancing.Authority == 'PJM',]
pjm_power = plot_power_demand_df(df)
pjm_power
```

#### Electric Reliability Council of Texas Authority
```{r viewing_the_power_dataset_erco_authority}
df = power_balance_data[power_balance_data$Balancing.Authority == 'ERCO',]
erco_power = plot_power_demand_df(df)
erco_power
```

#### Florida Power & Light Authority
```{r viewing_the_power_dataset_fpl_authority}
df = power_balance_data[power_balance_data$Balancing.Authority == 'FPL',]
fpl_power = plot_power_demand_df(df)
fpl_power
```

#### Midcontinent Independent System Operator Authority
```{r viewing_the_power_dataset_miso_authority}
df = power_balance_data[power_balance_data$Balancing.Authority == 'MISO',]
miso_power = plot_power_demand_df(df)
miso_power
```

Upon initial investigation, it is concerning that there is not a specific trend
for power demand across all of these major authorities. However, the region in
which these authorities operate cannot be forgotten. None of the selected
authorities operate within the same region and thus, each region must be
investigated independently when comparing them against ozone concentration.

## Looking at Ozone Concentration in the Atmosphere Above the United States during the Last six Months of 2019

The EPA has over 80 sites across the United States that monitor ozone
concentration within the atmosphere. Just as I did when looking at power
demand, I will only look at 5 of these sites. These sites will be located
in roughly the same region as the power authorities to allow for correlation
between power demand and ozone concentration to be accurately asessed. The
sites that will be analyzed are: LAV410, LRL117, ALC188, IRL141, and ALH157.

The sites IDs were obtained by using the
[CASTNET Site Locations Map](https://www.epa.gov/castnet/castnet-site-locations).

```{r plot_ozone_concentration_df}
# Generic function to simplify plotting of ozone concentration
plot_ozone_concentration_df <- function(df) {
  ggplot(data=df,aes(x=df$Date,y=as.numeric(df$OZONE_8HR_DAILY_MAX))) +
    geom_point() +
    xlab('Date') +
    ylab('Ozone Concentration (PPM)') +
    # Scale date by month to reduce number of labels.
    scale_x_discrete(breaks=c('07/01/2019','08/01/2019','09/01/2019','10/01/2019','11/01/2019','12/01/2019'))
}
```

#### LV410 (California Independent System Operator Authority)
```{r viewing_the_ozone_concentration_for_the_ciso_authority}
df = ozone_daily_max_data[ozone_daily_max_data$SITE_ID == 'LAV410',]
ciso_ozone = plot_ozone_concentration_df(df)
```

#### LRL117 (PJM Authority)
```{r viewing_the_ozone_concentration_for_the_pjm_authority}
df = ozone_daily_max_data[ozone_daily_max_data$SITE_ID == 'LRL117',]
pjm_ozone = plot_ozone_concentration_df(df)
pjm_ozone
```

#### ALC188 (Electric Reliability Council of Texas Authority)
```{r viewing_the_ozone_concentration_for_the_erco_authority}
df = ozone_daily_max_data[ozone_daily_max_data$SITE_ID == 'ALC188',]
erco_ozone = plot_ozone_concentration_df(df)
erco_ozone
```

#### IRL141 (Florida Power & Light Authority)
```{r viewing_the_ozone_concentration_for_the_fpl_authority}
df = ozone_daily_max_data[ozone_daily_max_data$SITE_ID == 'IRL141',]
fpl_ozone = plot_ozone_concentration_df(df)
fpl_ozone
```

#### ALH157 (Midcontinent Independent System Operator Authority)
```{r viewing_the_ozone_concentration_for_the_miso_authority}
df = ozone_daily_max_data[ozone_daily_max_data$SITE_ID == 'ALH157',]
miso_ozone = plot_ozone_concentration_df(df)
miso_ozone
```

Again, each region does not share a general trend, but both LRL117 and
ALH157 show a downwards trend in ozone concentration PPM of roughly
30-40%. Since both of these display a downwards trend, I will start by
investigating the power demand by the PJM and Midcontinent Independent
System Operator Authorities.

## Investigationg Both Datasets to Find Correlations

#### CISO Authority

```{r ciso_authority_review}
grid.arrange(
  ciso_power,
  ciso_ozone,
  ncol = 1,
  nrow = 2)
```

When looking at the power demand over time and the decrease in ozone
concentration over time, there is a potential correlation between the two
datasets. Towards the middle to end of September, the overall power demand and
ozone concentration in the CISO authority both see a decrease. Additionally, in
December, both the power demand and ozone concentration begin to rise again.
For the CISO authority area, California, there appears to be some correlation
between the two datasets.

#### PJM Authority

```{r pjm_authority_review}
grid.arrange(
  pjm_power,
  pjm_ozone,
  ncol = 1,
  nrow = 2)
```

When looking at the power demand over time and the decrease in ozone
concentration over time, there is no real correlation between the two. While
there is a major shift in power demand around the same time that the ozone
concentration begins to drop a couple of days into October, however, that is
the only similarity between the data. The power demand returns to a steady flow
towards the beginning of November while the ozone concentration continues to
drop. For the PJM authority area, the DMV, there is no increase or decrease of
power demand based on a change in ozone concentration.

#### ERCO Authority

```{r erco_authority_review}
grid.arrange(
  erco_power,
  erco_ozone,
  ncol = 1,
  nrow = 2)
```

When looking at the power demand over time and the decrease in ozone
concentration over time, there is no correlation between the two datasets.
While there is a decrease if power demand from the middle of October to the end
of the year, there is no major change in the ozone concentration for the ECRO
authority area, Texas.

#### FPL Authority

```{r fpl_authority_review}
grid.arrange(
  fpl_power,
  fpl_ozone,
  ncol = 1,
  nrow = 2)
```

When looking at the power demand over time and the decrease in ozone
concentration over time, there is some correlation between the two datasets.
As power demand increases in the late night to early hours of the morning, the
ozone concentration has also inceased. However, this is only for the early
morning hours and the overall mejority of power demand has decreased since
November.

#### MISO Authority

```{r miso_authority_review}
grid.arrange(
  miso_power,
  miso_ozone,
  ncol = 1,
  nrow = 2)
```

When looking at the power demand over time and the decrease in ozone
concentration over time, there is some correlation between the two datasets.
Around the middle to end of October, both graphs show an increase in variance
for both power demand and ozone concentration, which both continue on until
the end of December. However, the variance of power demand is not nearly as
dramatic as the variance and drop of ozone concentration.

## Summarizing Findings

Wrapping up this short investigation reveals that there may be a slight
correlation between the two datasets that could lead to further investigation.
However, there is not a large enough correlation to definitively say that
power demand and ozone concentration within the atmosphere of the United States
are correlated, or that one is a result of the other. There are obviously more
factors than just ozone concentration that result in an increase or decrease in
power demand, however, ozone concentration is likely not a factor at all.
Instead, further investigation as to the age of infatsturcture with the
authorities should conducted to investigate the efficiencies of the power
facilities that currently exists to see if power demand can be meet,
maintained, and decreased with more efficient and cheaper hardware.
